### 2017-12-06 Analysis/data_cleaning.Rmd
data cleaning pipeline

---
title: "Cleaning the data"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Sorting the disparity methods dataset

```{r}
source("../Functions/fuzzy.sort.data.R")
```

First lets load the data:

```{r}
## Reading the data
data_raw <- read.csv("../Data/disparity_methods.csv", stringsAsFactors = FALSE)

## Keeping only the studies with data
data <- data_raw[apply(data_raw[, 3:9], 1, function(X) ifelse(all(X == "-"), FALSE, TRUE)), ]

## Removing rogue NA rows
data <- data[-231,]
```

Then, lets sort the following variables:

 * The focal biological group
 * The type of data (3D landmarks, discrete data, etc...)
 * Whether the full organism was used or not.
 * How was the morphospace explicitly defined?
 * How was disparity explicitly defined?
 * Which test was applied for testing hypothesis?
 * Was phylogeny taken into account?

## Biological group

Sorting the biological groups into broader categories

```{r}
## Biological group
data_taxa <- data[,3]

## Mammals synonyms
mammals <- c("Felids", "Carnivora", "Primates", "Rodents", "Whales", "Cetacean", "Cetaceans", "Metatheria", "Marsupials", "Humans", "Perissodactyla", "Lipotyphla", "Mammals")
data_taxa <- ifelse(data_taxa %in% mammals, "mammals", data_taxa)

## Aves synonyms
aves <- c("Meliphagidae", "Birds")
data_taxa <- ifelse(data_taxa %in% aves, "aves", data_taxa)

## Archosaurs (non-aves)
archosaurs <- c("Pterosaurs", "Crocodiles", "Sauropoda", "Dinosaurs", "Theropods", "Archosauria")
data_taxa <- ifelse(data_taxa %in% archosaurs, "archosaurs", data_taxa)

## Squamates
squamates <- c("Squamates", "Rhynchocephalia", "Anolis", "Squamates")
data_taxa <- ifelse(data_taxa %in% squamates, "squamates", data_taxa)

## Other tetrapods
amphib_other_tetrapods <- c("Plesiosaur", "Captorhinidae", "Anomodonts", "Ichtyosaurs", "Amphibians", "Frogs", "Caecilian")
data_taxa <- ifelse(data_taxa %in% amphib_other_tetrapods, "amphib_other_tetrapods", data_taxa)

## Arthropods
arthropods <- c("Odonata", "Trilobites", "Ostracodes", "Scorpions", "Insects", "Eurypterida", "Orthoptera", "Arthropods", "Crustacean", "Beetles", "Hymenoptera", "Amphipodes")
data_taxa <- ifelse(data_taxa %in% arthropods, "arthropods", data_taxa)

## Plants
plantae <- c("Angiosperms", "Rubiacea", "Marchantiophyta", "Ferns + angiosperms")
data_taxa <- ifelse(data_taxa %in% plantae, "plantae", data_taxa)

## Echinoderms
echinoderms <- c("Echinoderms", "Crinoids", "Echinoids")
data_taxa <- ifelse(data_taxa %in% echinoderms, "echinoderms", data_taxa)

## Nematods
cnidaria_other_invert <- c("Nematode", "Brachiopods", "Onychophora", "Palaeoscolecid worms", "Graptolites", "Cnidera", "Corals", "Cnidaria")
data_taxa <- ifelse(data_taxa %in% cnidaria_other_invert, "cnidaria_other_invert", data_taxa)

## Actinopterigygians
actino_chondry <- c("Cichlids", "Fish", "Fishes", "Shark")
data_taxa <- ifelse(data_taxa %in% actino_chondry, "actino_chondry", data_taxa)

## Molluscs
molluscs <- c("Ammonites", "Gastropod", "Ammonoids", "Cephalopoda", "Mollusca", "Bivalves")
data_taxa <- ifelse(data_taxa %in% molluscs, "molluscs", data_taxa)

## Unicellulars
unicellulars <- c("Foraminifera", "Desimidiales", "Acritarch")
data_taxa <- ifelse(data_taxa %in% unicellulars, "unicellulars", data_taxa)

## General
general <- c("Metazoans", "Ediacarian", "Blastozoa + Crinoids + Trilobite", "Echinoids + Trilobites", "Crinoids + Balstozoa", "Marine invertebrates", "Amniotes", "Vertebrates")
data_taxa <- ifelse(data_taxa %in% general, "general", data_taxa)

## Simulated
simualted <- c("Simulated", "-", "")
data_taxa <- ifelse(data_taxa %in% simualted, "simualted", data_taxa)

## Selecting the categories
data_taxa_categories <- unique(data_taxa)

data[,3] <- data_taxa
```

## Type of data

```{r}
## Types of data
data_type <- data[,4]

unique(data_type)

## continuous synonyms
continuous <- c("1D Continuous", "DTT pipeline", "Continuous")
data_type <- ifelse(data_type %in% continuous, "continuous", data_type)

## Math transfo
math_transfo <- c("B-spline", "Elliptic Fourier", "Outlines", "Ellipctic Fourier")
data_type <- ifelse(data_type %in% math_transfo, "math_transfo", data_type)

## Other_univar
other_univar <- c("Body mass", "Metabolic Rate", "Body size", "Glands for chemical communication", "Chemical signals", "Colours")
data_type <- ifelse(data_type %in% other_univar, "other_univar", data_type)

## Discrete
data_type <- ifelse(data_type == "Discrete", "discrete", data_type)

## Non
none <- c(NA, "-")
data_type <- ifelse(data_type %in% none, "none", data_type)

## 2d + discrete
d2_discrete <- c("Discrete + 2D landmarks")
data_type <- ifelse(data_type %in% d2_discrete, "discrete + 2D landmarks", data_type)

## 2d + continuous
d2_continu <- c("Continuous + 2D landmarks")
data_type <- ifelse(data_type %in% d2_continu, "continuous + 2D landmarks", data_type)

## continuous + other
continu_other <- c("Continuous + Body size")
data_type <- ifelse(data_type %in% continu_other, "continuous + other_univar", data_type)

## continuous + discrete
continu_discrete <- c("Continuous + discrete" )
data_type <- ifelse(data_type %in% continu_discrete, "continuous + discrete", data_type)

## Selecting the categories
data_type_categories <- unique(unlist(strsplit(unique(data_type), split = " \\+ ")))

data[,4] <- data_type
```

## Full organism

```{r}
## Full organism
full_organism <- data[,5]

## Selecting the categories
full_organism_categories <- c("Yes", "No", NA)
```

## Morphospaces

```{r}
## Morphospaces
morphospaces <- data[,6]

unique(morphospaces)

## PCA
PCA <- c("PCA", "Procrustes PCA", "pPCA", "Regression PCA")
morphospaces <- ifelse(morphospaces %in% PCA, "PCA", morphospaces)

## PCO PCoA
PCO <- c("PCO", "PCoA", "PcoA", "FMD (Foote 1993)")
morphospaces <- ifelse(morphospaces %in% PCO, "PCO", morphospaces)

## MDS NMDS
MDS <- c("MDS", "NMDS", "nMDS")
morphospaces <- ifelse(morphospaces %in% MDS, "MDS", morphospaces)

## RWA
RWA <- c("RWA", "RWA (Relative Warp Analysis)")
morphospaces <- ifelse(morphospaces %in% RWA, "RWA", morphospaces)

## Canonical Analysis
canonical <- c("Canonical variables", "canonical correlation analysis (CCA)")
morphospaces <- ifelse(morphospaces %in% canonical, "canonical", morphospaces)

## Correspondence Analysis
correspondence <- c("Correspondence analysis", "MCA (multiple correspondence analyses)")
morphospaces <- ifelse(morphospaces %in% correspondence, "correspondence", morphospaces)

## Other
other <- c("Eigen shape analysis", "Factorial planes","Biplot (two traits only)", "biplot (residuals of two traits)", "Factorial planes","Procrustes plane (2 D axis)", "Procrustes coordinates")
morphospaces <- ifelse(morphospaces %in% other, "other", morphospaces)

## None
none <- c("DTT pipeline")
morphospaces <- ifelse(morphospaces %in% none, "none", morphospaces)
morphospaces <- ifelse(is.na(morphospaces), "none", morphospaces)

## Combos
morphospaces <- ifelse(morphospaces %in% "PCA + canonical variables", "PCA + canonical", morphospaces)
morphospaces <- ifelse(morphospaces %in% "Dissimilarity matric + PCoA", "PCO + other", morphospaces)
morphospaces <- ifelse(morphospaces %in% "RWA (Relative Warp Analysis) + PCO", "PCO + RWA", morphospaces)
morphospaces <- ifelse(morphospaces %in% "CVA + PCA" , "PCA + canonical", morphospaces)

## Selecting the categories
morphospaces_categories <- unique(unlist(strsplit(unique(morphospaces), split = " \\+ ")))

```

```{r}
disparity <- c("sum of variance", "sum of range", "product of variance", "product of range", "distance", "observed") #TG: Area from observed to estimated (DTT)
test <- c("anova", "model fitting", "canonical", "confidence interval")
phylo <- c("yes", "no", "mapped")

factors <- list("data type" = data_type, "full organism" = full_organism, "morphospace" = morphospace, "disparity" = disparity, "test" = test, "phylo correction" = phylo)
```