### 2017-12-08 Analysis/data_cleaning.Rmd
cleaned data

---
title: "Cleaning the data"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Sorting the disparity methods dataset

```{r}
source("../Functions/fuzzy.sort.data.R")
```

First lets load the data:

```{r}
## Reading the data
data_raw <- read.csv("../Data/disparity_methods.csv", stringsAsFactors = FALSE)

## Keeping only the studies with data
data <- data_raw[apply(data_raw[, 3:9], 1, function(X) ifelse(all(X == "-"), FALSE, TRUE)), ]
```

Then, lets sort the following variables:

 * The focal biological group
 * The type of data (3D landmarks, discrete data, etc...)
 * Whether the full organism was used or not.
 * How was the morphospace explicitly defined?
 * How was disparity explicitly defined?
 * Which test was applied for testing hypothesis?
 * Was phylogeny taken into account?

## Biological group

Sorting the biological groups into broader categories

```{r}
## Biological group
data_taxa <- data[,3]

## Mammals synonyms
mammals <- c("Felids", "Carnivora", "Primates", "Rodents", "Whales", "Cetaceans", "Metatheria", "Marsupials", "Humans", "Perissodactyla", "Lipotyphla", "Mammals", "mammals")
data_taxa <- multi.gsub(mammals, "mammals", data_taxa)

## Aves synonyms
aves <- c("Meliphagidae", "Birds")
data_taxa <- multi.gsub(aves, "aves", data_taxa)

## Archosaurs (non-aves)
archosaurs <- c("Pterosaurs", "Crocodiles", "Sauropoda", "Dinosaurs", "Theropods", "Archosauria")
data_taxa <- multi.gsub(archosaurs, "archosaurs", data_taxa)

## Squamates
squamates <- c("Squamates", "Rhynchocephalia", "Anolis", "Squamates")
data_taxa <- multi.gsub(squamates, "squamates", data_taxa)

## Other tetrapods
amphib_other_tetrapods <- c("Plesiosaur", "Captorhinidae", "Anomodonts", "Ichtyosaurs", "Amphibians", "Frogs", "Caecilian")
data_taxa <- multi.gsub(amphib_other_tetrapods, "amphib_other_tetrapods", data_taxa)

## Arthropods
arthropods <- c("Odonata", "Trilobites", "Trilobite", "Ostracodes", "Scorpions", "Insects", "Eurypterida", "Orthoptera", "Arthropods", "Crustacean", "Beetles", "Hymenoptera", "Amphipodes")
data_taxa <- multi.gsub(arthropods, "arthropods", data_taxa)

## Plants
plantae <- c("Angiosperms", "Rubiacea", "Marchantiophyta", "angiosperms", "Ferns")
data_taxa <- multi.gsub(plantae, "plantae", data_taxa)

## Echinoderms
echinoderms <- c("Echinoderms", "Crinoids", "Echinoids", "Blastozoa", "Balstozoa")
data_taxa <- multi.gsub(echinoderms, "echinoderms", data_taxa)

## Nematods
other_invert <- c("Nematode", "Brachiopods", "Onychophora", "Palaeoscolecid worms", "Graptolites", "Cnidera", "Corals", "Cnidaria")
data_taxa <- multi.gsub(other_invert, "other_invert", data_taxa)

## Actinopterigygians
actino_chondry <- c("Cichlids", "Fish", "Fishes", "Shark")
data_taxa <- multi.gsub(actino_chondry, "actino_chondry", data_taxa)

## Molluscs
molluscs <- c("Ammonites", "Gastropod", "Ammonoids", "Cephalopoda", "Mollusca", "Bivalves")
data_taxa <- multi.gsub(molluscs, "molluscs", data_taxa)

## Unicellulars
unicellulars <- c("Foraminifera", "Desimidiales", "Acritarch")
data_taxa <- multi.gsub(unicellulars, "unicellulars", data_taxa)

## General
general <- c("Metazoans", "Ediacarian", "Marine invertebrates", "Amniotes", "Vertebrates", "Marine ingeneral")
data_taxa <- multi.gsub(general, "general", data_taxa)

## Simulated
simulated <- c("Simulated", "-", "")
data_taxa <- ifelse(data_taxa %in% simulated, "simulated", data_taxa)

## Selecting the categories
data_taxa_categories <- unique(unlist(strsplit(unique(data_taxa), split = " \\+ ")))

data[,3] <- data_taxa
```

## Type of data

```{r}
## Types of data
data_type <- data[,4]

## continuous synonyms
continuous <- c("1D Continuous", "DTT pipeline", "Continuous")
data_type <- multi.gsub(continuous, "continuous", data_type)

## Math transfo
math_transfo <- c("B-spline", "Elliptic Fourier", "Outlines", "Ellipctic Fourier")
data_type <- multi.gsub(math_transfo, "math_transfo", data_type)

## Other_univar
other_univar <- c("Body mass", "Metabolic Rate", "Body size", "Glands for chemical communication", "Chemical signals", "Colours")
data_type <- multi.gsub(other_univar, "other_univar", data_type)

## Discrete
data_type <- multi.gsub("Discrete", "discrete", data_type)

data_type <- ifelse(data_type == "Discrete", "discrete", data_type)

## Non
none <- c(NA, "-")
data_type <- ifelse(data_type %in% none, "none", data_type)

## Selecting the categories
data_type_categories <- unique(unlist(strsplit(unique(data_type), split = " \\+ ")))

data[,4] <- data_type
```

## Full organism

```{r}
## Full organism
full_organism <- data[,5]

## Selecting the categories
full_organism_categories <- c("Yes", "No", NA)
```

## Morphospaces

```{r}
## Morphospaces
morphospaces <- data[,6]

## PCA
PCA <- c("Procrustes PCA", "pPCA", "Regression PCA")
morphospaces <- multi.gsub(PCA, "PCA", morphospaces)

## PCO PCoA
PCO <- c("PCoA", "PcoA", "FMD (Foote 1993)")
morphospaces <- multi.gsub(PCO, "PCO", morphospaces)

## MDS NMDS
MDS <- c("NMDS", "nMDS")
morphospaces <- multi.gsub(MDS, "MDS", morphospaces)

## RWA
RWA <- c("RWA\\ \\(Relative Warp Analysis\\)")
morphospaces <- multi.gsub(RWA, "RWA", morphospaces)

## Canonical Analysis
canonical <- c("Canonical variables", "canonical correlation analysis\\ \\(CCA\\)", "canonical variables", "CVA")
morphospaces <- multi.gsub(canonical, "canonical", morphospaces)

## Raup spce
morphospaces <- multi.gsub("Raup-space", "raup_space", morphospaces)

## Correspondence Analysis
correspondence <- c("Correspondence analysis", "MCA\\ \\(multiple correspondence analyses\\)")
morphospaces <- multi.gsub(correspondence, "correspondence", morphospaces)

## Other
other <- c("Eigen shape analysis", "Factorial planes", "Biplot\\ \\(two traits only\\)", "biplot\\ \\(residuals of two traits\\)", "Procrustes plane \\(2 D axis\\)", "Procrustes coordinates", "FMD \\(Foote 1993\\)", "Dissimilarity matric")
morphospaces <- multi.gsub(other, "other", morphospaces)

## None
none <- c("DTT pipeline", "-")
morphospaces <- ifelse(morphospaces %in% none, "none", morphospaces)
morphospaces <- ifelse(is.na(morphospaces), "none", morphospaces)

## Selecting the categories
morphospaces_categories <- unique(unlist(strsplit(unique(morphospaces), split = " \\+ ")))

data[,6] <- morphospaces
```

## Disparity

```{r}
## Disparity
disparity <- data[,7]

number_unique_disparities <- length(unique(tolower(disparity)))

## variances
variance <- c("PCA variance", "sum of variance", "Sum of variance", "product of variances", "sum of variances", "Variance", "mean of variance", "sum of variances", "Total variance", "nth root of variance", "product of variance", "geometric mean of variance", "Procrustes variance", "Product of variance", "root product of variances", "total variance", "nth root of varaince")
disparity <- multi.gsub(variance, "variance", disparity)

## Sum of range
range <- c("sum of ranges", "product of range", "product of ranges", "mean of range", "range", "total range", "sum of range", "nth root of range", "Range", "geometric mean of range", "root product of ranges", "mean or range")
disparity <- multi.gsub(range, "range", disparity)

## Distance
distances <- c("Gower distance", "Euclidean distance", "mean square euclidean distance", "Procrustes distances \\(geomorph\\)", "other distances", "Procrustes distances", "Average square distance", "mean distance", "mean euclidean distance", "mean Manhattan distance", "Euclidean distance between groups", "average squared distance", "mean euclidean distance", "Procrustes distance", "MNND \\(Mean nearest neighbour distance\\)", "Euclidean Procrustes distances", "sum of Procrustes distances", "range of maximum euclidean distance", "average Procrustes distance", "sum of the squared Procrustes distances")
disparity <- multi.gsub(distances, "distances", disparity)

## Pairwise distance
pair_distances <- c("DTT \\(morpho evolution in relation to diversification\\)", "DTT pairwise Euclidean distance", "weighted mean pairwise dissimilarity", "pairwise euclidean distance", "DTT pairwise average squared Euclidean distance", "mean pairwise dissimilarity", "sum of squared pairwise distance between coordinates", " pairwise distance", "average pairwise disparity", "Average pairwise squared distance", "average pairwise dissimilarity", "mean square distance between sp", "distance between species", "pairwise character dissimilarity", "mean pairwise distance", "pairwise distance", "mean squared pairwise difference in body size", "average morphological dissimilarity", "Average pairwise dissimilarity", "Pairwise Mahalanobis distance", "sum of the absolute distance for each variable", "DTT pairwise squared distances")
disparity <- multi.gsub(pair_distances, "pair_distances", disparity)

## Distance from centroid
centroids <- c("distance from centroids", "centroid", "Procrustes distances from centre", "distance from centroids", "distance from neighbour", "furthest distance", "Average square distance from centroid", "distance from centroid", "centroid size", "distance from centroid", "Group centroid", "mean distance from centroids", "mean distance to centroid", "mean Procrustes distance from centroid", "distances to centroids")
disparity <- multi.gsub(centroids, "centroids", disparity)

## Other
other <- c("MDI \\(Morphological disparity index - observed disparity vs median of expected from BM\\)", "Area between observed and median", "Foote's disparity measurement \\(Foote 1993, Neige 2003\\)", "participation ratio", "Body mass", "Regression score", "lineage density", "function of distance to ancestor and character difference")
disparity <- multi.gsub(other, "other", disparity)

## Character based
character <- c("Number of polymorphic characters", "number of unique pairwise characters", "unique pairwise combinations", "mean character difference")
disparity <- multi.gsub(character, "character", disparity)

## PC scores
ord_scores <- c("PC1\\&2 values", "space coordinates", "PC contrast values", "sum of the diagonal of the variance covariance matrix of the PCA", "Probability of encountering neighbouring data points in 3D")
disparity <- multi.gsub(ord_scores, "ord_scores", disparity)

## Volume/surface
volume <- c("Volume", "convex hull surface", "volume", "convex hull volume", "convex hull", "alpha shape volume", "Disparity Box 3D", "Number of cubes of theoretical morphospace", "PC scores", "PC axis scores")
disparity <- multi.gsub(volume, "volume", disparity)

## Missing
disparity <- ifelse(disparity == "?", NA, disparity)
disparity <- ifelse(disparity == "-", NA, disparity)

## Clean the matchings
disparity <- gsub("ss", "s", gsub("ss", "s", disparity))
disparity <- gsub(" and ", " \\+ ", disparity)
disparity <- gsub("DTT pair_distances", "pair_distances", disparity)
disparity <- gsub("DTT pair_distances", "pair_distances", disparity)
disparity <- gsub("range variance", "range \\+ variance", disparity)
disparity <- gsub("distances from centre", "centroids", disparity)
disparity <- gsub("distances from centroids", "centroids", disparity)
disparity <- gsub("distances to centroids", "centroids", disparity)
disparity <- gsub("mean ", "", disparity)
disparity <- gsub("median ", "", disparity)
disparity <- gsub("average ", "", disparity)
disparity <- gsub("other _ character", "other + character", disparity)
disparity <- gsub("range \\+variance", "range \\+ variance", disparity)
disparity <- gsub("geometric ", "", disparity)

disparity_categories <- unique(unlist(strsplit(unique(disparity), split = " \\+ ")))

data[,7] <- disparity
```

## Tests

```{r}
## Disparity
tests <- data[,8]

## ANOVA
anova <- c("MANOVA", "PERMANOVA", "ANOVA", "NPMANOVA", "Procrustes ANOVA", "phyloANOVA", "PERMANOVA", "ANCOVA", "phyloANOVA", "Kruskal-Wallis test", "coefficient of variance", "permutation test", "Anova", "Ripley's K", "MANCOVA", "Chi-squared")
tests <- multi.gsub(anova, "anova", tests)

## LDA
LDA <- c("LDA", "phyloLDA")
tests <- multi.gsub(LDA, "LDA", tests)

## Model test
models <- c("Model fitting", "comparison to a null expectancy", "bootstrap/rarefaction null hypothesis")
tests <- multi.gsub(models, "models", tests)

## Mean test
t_test <- c("t-test", "Mann-Whitney", "Walsh F-test", "Z-test", "F-statistics", "Wilcoxon", "multivariate t-test \\(Hotellings' T\\^2 test\\)")
tests <- multi.gsub(t_test, "t_test", tests)

## Correlation
correlation <- c("Correlation \\(Fisher's exact test \\+ Spearman's rank\\)", "Mantel test", "correlation tests", "Spearman correlation")
tests <- multi.gsub(correlation, "correlation", tests)

## regressions
regressions <- c("PGLS \\(PC \\~ trait\\)", "Partial Least Squares", "Canonical analysis \\(PCM style\\)", "lm", "Covariance ratio", "GLS", "distance PGLS", "multivariate regression", "linear regressions", "GLS multiple regression", "PICs", "Multivariate PGLS", "Non-linear regressions", "regression score", "Breush-Pagan")
tests <- multi.gsub(regressions, "regressions", tests)

## non stat
non_stat <- c("CI overlap", "confidence interval overlap")
tests <- multi.gsub(non_stat, "non_stat", tests)

## other
other <- c("Morphological Disparity Index \\(MDI\\) statistics", "centre of gravity", "Canonical variates analysis", "Canonical variance analysis", "Multidimensional Convergence Index", "Kernel analysis", "disparity test")
tests <- multi.gsub(other, "other", tests)

## clustering
clustering <- c("Nearest-neighbour test", "Cluster dendrogram analysis", "neighbour joining of the centroids", "Nearest-neighbour clustering", "Hierarchical clustering", "Branching random walks")
tests <- multi.gsub(clustering, "clustering", tests)

tests <- ifelse(tests == "", NA, tests)
tests <- ifelse(tests == "?", NA, tests)
tests <- ifelse(tests == "-", NA, tests)
tests <- ifelse(tests == "Raup-space", NA, tests)

## Clean the matchings
tests <- gsub("distance Plinear", "linear", tests)
tests <- gsub("multivariate t_test \\(Hotellings' T\\^2 test\\)", "t_test", tests)
tests <- gsub("Kruskal-Wallis", "anova", tests)
tests <- gsub("distance Pregressions", "regressions", tests)

tests_categories <- unique(unlist(strsplit(unique(tests), split = " \\+ ")))

data[,8] <- tests
```

## Phylogeny

```{r}
phylogeny <- data[,9]

phylogeny <- multi.gsub("Yes \\(pPC\\)", "pPC", phylogeny)
phylogeny <- ifelse(phylogeny == "", NA, phylogeny)

phylogeny_categories <- unique(unlist(strsplit(unique(phylogeny), split = " \\+ ")))

data[,9] <- phylogeny
```

## Wrapping up the data set

```{r}
## Cleaned data
data <- data[,-c(1,2)]
cleaned_data <- list("data" = data)
cleaned_data$factors <- list(data_taxa = data_taxa_categories, data_type = data_type_categories, full_organism = full_organism_categories, morphospaces = morphospaces_categories, disparity = disparity_categories, tests = tests_categories, phylogeny = phylogeny_categories)
cleaned_data$unique_disparity <- number_unique_disparities

## Save the file
saveRDS(cleaned_data, file = paste0("../Data/Processed/cleaned_data.Rda"))

```