### 2017-12-18 Analysis/effect_dimensionality.Rmd
Dimensionality test

---
title: "Effect of dimensionality on metrics"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Effect of dimensionality on disparity metrics

## Preparing the data

```{r}
set.seed(123)
library(dispRity)
data(BeckLee_mat50)

## Measuring for different metrics
sum_range <- sapply(2:48, function(X) dispRity(BeckLee_mat50, metric = c(sum, ranges), dimensions = X), simplify = FALSE)
prod_range <- sapply(2:48, function(X) dispRity(BeckLee_mat50, metric = c(prod, ranges), dimensions = X), simplify = FALSE)
sum_var <- sapply(2:48, function(X) dispRity(BeckLee_mat50, metric = c(sum, variances), dimensions = X), simplify = FALSE)
prod_var <- sapply(2:48, function(X) dispRity(BeckLee_mat50, metric = c(prod, variances), dimensions = X), simplify = FALSE)
med_centroid <- sapply(2:48, function(X) dispRity(BeckLee_mat50, metric = c(median, centroids), dimensions = X), simplify = FALSE)
avg_sq <- sapply(2:48, function(X) dispRity(BeckLee_mat50, metric = function(matrix){mean(pairwise.dist(matrix)^2)}, dimensions = X), simplify = FALSE)
ell_vol <- sapply(2:48, function(X) dispRity(BeckLee_mat50, metric = ellipse.volume, dimensions = X), simplify = FALSE)
n_ball <- sapply(2:48, function(X) dispRity(BeckLee_mat50, metric = n.ball.volume, dimensions = X), simplify = FALSE)

## Extracting the observed values
sum_range <- unlist(lapply(sum_range, function(X) return(summary(X)$obs)))
prod_range <- unlist(lapply(prod_range, function(X) return(summary(X)$obs)))
sum_var <- unlist(lapply(sum_var, function(X) return(summary(X)$obs)))
prod_var <- unlist(lapply(prod_var, function(X) return(summary(X)$obs)))
med_centroid <- unlist(lapply(med_centroid, function(X) return(summary(X)$obs)))
avg_sq <- unlist(lapply(avg_sq, function(X) return(summary(X)$obs)))
ell_vol <- unlist(lapply(ell_vol, function(X) return(summary(X)$obs)))
n_ball <- unlist(lapply(n_ball, function(X) return(summary(X)$obs)))

## Scale these scores
scale.max <- function(x) return(x/max(x))
sum_range_sca <- scale.max(sum_range)
prod_range_sca <- scale.max(prod_range)
sum_var_sca <- scale.max(sum_var)
prod_var_sca <- scale.max(prod_var)
med_centroid_sca <- scale.max(med_centroid)
avg_sq_sca <- scale.max(avg_sq)
ell_vol_sca <- scale.max(ell_vol)
n_ball_sca <- scale.max(n_ball)

metrics <- list(sum_range_sca, prod_range_sca, sum_var_sca, prod_var_sca, med_centroid_sca, avg_sq_sca, ell_vol_sca, n_ball_sca)
metrics_names <-  c("sum of ranges", "product of ranges", "sum of variance", "product of variance", "median distance from centroid", "average squared pairwise distance", "ellispoid volume", "minimum n-ball volume")
```

```{r, fig.width = 12, fig.height = 12}
## gg plot colours
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

cols <- gg_color_hue(8)
ylim <- c(0,1)
xlim <- c(2,48)

## Plotting the results
pdf("../Writing/Figures/DimensionsEffect.pdf", width = 12, height = 12)
op <- par(bty = "n")

plot(1,1, type = "n", ylim = ylim, xlim = xlim, xlab = "Dimensions", ylab = "Relative disparity")

for(metric in 1:8) {
    lines(x = c(2:48), metrics[[metric]], col = cols[metric])
}

legend("right", legend = metrics_names, col = cols, lty = 1)
par(op)
dev.off()
```