### 2018-02-27 Writing/Figures/Figures_in_R.Rmd
Overlap figure

---
title: "Figures"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Methods part

## Visualisation

A multidimensional space similar to Figure 3 in Davis et al 2012 Nature.

```{r}
set.seed(1)
## Sampling the four different groups
size <- 10            # x                      #y                       #z
group1 <- cbind(rnorm(size, -0.5, 0.25), rnorm(size,  0.5, 0.25), rnorm(size,  0.5, 0.25))
group2 <- cbind(rnorm(size, -0.5, 0.25), rnorm(size,  0.5, 0.25), rnorm(size, -0.5, 0.25))
group3 <- cbind(rnorm(size,  0.5, 0.25), rnorm(size,  0.5, 0.25), rnorm(size,  0.5, 0.25))
group4 <- cbind(rnorm(size,  0.5, 0.25), rnorm(size, -0.5, 0.25), rnorm(size,  0.5, 0.25))
groups <- list(group1, group2, group3, group4)

## Chull coordinates function
get.chull.coords <- function(x, axis) {
    chull_points <- chull(x[, c(axis)])
    chull_coords <- rbind(x[chull_points, axis], x[chull_points[1], axis])
}

## Getting the coordinates
chull_coords_xy <- lapply(groups, get.chull.coords, axis = c(1, 2))
chull_coords_xz <- lapply(groups, get.chull.coords, axis = c(1, 3))

## Empty plot function
empty.plot <- function(xlab = "Dimension 1", ylab) {
    par(bty = "n")
    plot(NULL, xlim = c(-1.1, 1.1), ylim = c(-1.1, 1.1), xlab = xlab, ylab = ylab, xaxt='n', yaxt='n')
    abline(v = 0)
    abline(h = 0)
}

## Setting the colour palette
col_palette <- as.list(viridis::viridis(4))

par(mfrow = c(2, 1))
par(mar = c(1, 4, 1, 3))
## Plotting the first coordinates sets
empty.plot(xlab = "", ylab = "Dimension 2")
text(-1,1, "1", cex = 2)
## Adding the points
plot.points <- function(x, col, axis) points(x[, axis], col = col, pch = 19, cex = 0.7)
silent <- mapply(plot.points, groups, col_palette, MoreArgs = list(axis = c(1, 2)))
## Adding the polygons
plot.polygons <- function(x, col) polygon(x[, 1], x[, 2], col = col, density = 10, border = col)
silent <- mapply(plot.polygons, chull_coords_xy, col_palette)
legend("bottomlef", legend = LETTERS[1:4], col = unlist(col_palette), bty = "n", pch = 19, lty = 1)

par(mar = c(4, 4, 1, 3))
## Plotting the second coordinates sets
empty.plot(ylab = "Dimension 3")
text(-1,1, "2", cex = 2)
## Adding the points
silent <- mapply(plot.points, groups, col_palette, MoreArgs = list(axis = c(1, 3)))
## Adding the polygons
silent <- mapply(plot.polygons, chull_coords_xz, col_palette)

```

```{r, eval = FALSE, message = FALSE, fig.width = 12, fig.height = 12, echo = FALSE}
```