### 2018-11-02 Analysis/Figures_in_R.Rmd
Moved figures Rmd to analysis

---
title: "Figures"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Methods part

```{r}
library(dispRity)
```

## Visualisation

A multidimensional space similar to Figure 3 in Davis et al 2012 Nature.

```{r, fig.width = 5, fig.height = 8.33333}
set.seed(42)
## Sampling the four different groups
size <- 10            # x                      #y                       #z
group1 <- cbind(rnorm(size, -0.5, 0.25), rnorm(size,  0.5, 0.25), rnorm(size,  0.5, 0.25))
group2 <- cbind(rnorm(size, -0.5, 0.25), rnorm(size,  0.5, 0.25), rnorm(size, -0.5, 0.25))
group3 <- cbind(rnorm(size,  0.5, 0.25), rnorm(size,  0.5, 0.25), rnorm(size,  0.5, 0.25))
group4 <- cbind(rnorm(size,  0.5, 0.25), rnorm(size, -0.5, 0.25), rnorm(size,  0.5, 0.25))
groups <- list(group1, group2, group3, group4)

## Chull coordinates function
get.chull.coords <- function(x, axis) {
    chull_points <- chull(x[, c(axis)])
    chull_coords <- rbind(x[chull_points, axis], x[chull_points[1], axis])
}

## Getting the coordinates
chull_coords_xy <- lapply(groups, get.chull.coords, axis = c(1, 2))
chull_coords_xz <- lapply(groups, get.chull.coords, axis = c(1, 3))

## Empty plot function
empty.plot <- function(xlab = "Dimension 1", ylab) {
    par(bty = "n")
    plot(NULL, xlim = c(-1.1, 1.1), ylim = c(-1.1, 1.1), xlab = xlab, ylab = ylab, xaxt='n', yaxt='n')
    abline(v = 0)
    abline(h = 0)
}

## Setting the colour palette
col_palette <- as.list(viridis::viridis(4))

plot.figure.dimensionality <- function(){
    par(mfrow = c(2, 1))
    par(mar = c(1, 4, 1, 3))
    ## Plotting the first coordinates sets
    empty.plot(xlab = "", ylab = "Dimension 2")
    text(-1,1, "1", cex = 2)
    ## Adding the points
    plot.points <- function(x, col, axis) points(x[, axis], col = col, pch = 19, cex = 0.7)
    silent <- mapply(plot.points, groups, col_palette, MoreArgs = list(axis = c(1, 2)))
    ## Adding the polygons
    plot.polygons <- function(x, col) polygon(x[, 1], x[, 2], col = col, density = 10, border = col)
    silent <- mapply(plot.polygons, chull_coords_xy, col_palette)
    legend("bottomlef", legend = LETTERS[1:4], col = unlist(col_palette), bty = "n", pch = 19, lty = 1)

    par(mar = c(4, 4, 1, 3))
    ## Plotting the second coordinates sets
    empty.plot(ylab = "Dimension 3")
    text(-1,1, "2", cex = 2)
    ## Adding the points
    silent <- mapply(plot.points, groups, col_palette, MoreArgs = list(axis = c(1, 3)))
    ## Adding the polygons
    silent <- mapply(plot.polygons, chull_coords_xz, col_palette)
}

plot.figure.dimensionality()

## Figure for the paper
pdf(file = "../Writing/Figures/figure_dimensionality.pdf", width = 5, height = 8.33333)
plot.figure.dimensionality()
dev.off()
```

## Metrics

```{r, fig.width = 10, fig.height = 10}
set.seed(42)

## Scale the results to be between -1 and 1
scale.matrix <- function(matrix) {
    ## Get the range for scaling
    range <- apply(matrix, 2, range)

    ## Shift to positive
    shift_matrix <- matrix
    shift_matrix[,1] <- matrix[,1] + abs(range[1,1])
    shift_matrix[,2] <- matrix[,2] + abs(range[1,2])

    ## Scale the matrix (to + 2)
    shift_matrix <- apply(shift_matrix, 2, function(x) x/(max(x)/2))

    ## Reshift the matrix
    return(shift_matrix - 1)
}

## Plotting the two matrices
plot.both.matrices <- function(matrix1, matrix2, col = c("blue", "orange"), ...) {
    plot(matrix1, pch = 21, bg = col[1], ..., xlab = "", ylab = "")
    points(matrix2, pch = 21, bg = col[2])
}

## Two similar correlation matrices
cor_mat_positive <- matrix(c(1, 0.8, 0.8, 1), 2, 2)
cor_mat_negative <- matrix(c(1, -0.8, -0.8, 1), 2, 2)

## Correlation differences
correlated_positive <- scale.matrix(space.maker(100, 2, rnorm, cor.matrix = cor_mat_positive))
correlated_negative <- scale.matrix(space.maker(100, 2, rnorm, cor.matrix = cor_mat_negative))

## Doughnut differences
outer <- space.maker(100, 2, random.circle, arguments = list(list(runif, inner = 0.5)))
inner <- space.maker(100, 2, random.circle, arguments = list(list(runif, outer = 0.5)))

## Density differences
base_matrix <- space.maker(100, 2, runif, arguments = list(list(max = 1, min = -1)))
modif_matrix <- apply(base_matrix, 2, function(x) x+rnorm(length(x), sd = 0.035))

## Position differences
upper_right <- space.maker(100, 2, rnorm, arguments = list(list(mean = 0.5, sd = 0.2)))
lower_left <- space.maker(100, 2, rnorm, arguments = list(list(mean = -0.5, sd = 0.2)))

plot.figure.metrics <- function(){
    par(mfrow = c(2,2), mar = c(2.5,2.5,2.5,2.5))
    plot.both.matrices(correlated_positive, correlated_negative, main = "Correlation")
    plot.both.matrices(outer, inner, main = "Filling")
    plot.both.matrices(base_matrix, modif_matrix, main = "Density")
    plot.both.matrices(upper_right, lower_left, xlim = c(-1, 1), ylim = c(-1, 1), main = "Position")
}

plot.figure.metrics()

## Figure for the paper
pdf(file = "../Writing/Figures/figure_metrics.pdf", width = 10, height = 10)
plot.figure.metrics()
dev.off()

```