### 2019-10-30 Analysis/Figures_in_R.Rmd
Updated figures

---
title: "Figures"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Methods part

```{r}
library(dispRity)
```

## Visualisation

A multidimensional space similar to Figure 3 in Davis et al 2012 Nature.

```{r, fig.width = 5, fig.height = 8.33333}
set.seed(42)
## Sampling the four different groups
size <- 10            # x                      #y                       #z
group1 <- cbind(rnorm(size, -0.5, 0.25), rnorm(size,  0.5, 0.25), rnorm(size,  0.5, 0.25))
group2 <- cbind(rnorm(size, -0.5, 0.25), rnorm(size,  0.5, 0.25), rnorm(size, -0.5, 0.25))
group3 <- cbind(rnorm(size,  0.5, 0.25), rnorm(size,  0.5, 0.25), rnorm(size,  0.5, 0.25))
group4 <- cbind(rnorm(size,  0.5, 0.25), rnorm(size, -0.5, 0.25), rnorm(size,  0.5, 0.25))
groups <- list(group1, group2, group3, group4)

## Chull coordinates function
get.chull.coords <- function(x, axis) {
    chull_points <- chull(x[, c(axis)])
    chull_coords <- rbind(x[chull_points, axis], x[chull_points[1], axis])
}

## Getting the coordinates
chull_coords_xy <- lapply(groups, get.chull.coords, axis = c(1, 2))
chull_coords_xz <- lapply(groups, get.chull.coords, axis = c(1, 3))

## Empty plot function
empty.plot <- function(xlab = "Dimension 1", ylab) {
    par(bty = "n")
    plot(NULL, xlim = c(-1.1, 1.1), ylim = c(-1.1, 1.1), xlab = xlab, ylab = ylab, xaxt='n', yaxt='n')
    abline(v = 0)
    abline(h = 0)
}

## Setting the colour palette
col_palette <- as.list(viridis::viridis(4))

plot.figure.dimensionality <- function(){
    par(mfrow = c(2, 1))
    par(mar = c(1, 4, 1, 3))
    ## Plotting the first coordinates sets
    empty.plot(xlab = "", ylab = "Dimension 2")
    text(-1,1, "1", cex = 2)
    ## Adding the points
    plot.points <- function(x, col, axis) points(x[, axis], col = col, pch = 19, cex = 0.7)
    silent <- mapply(plot.points, groups, col_palette, MoreArgs = list(axis = c(1, 2)))
    ## Adding the polygons
    plot.polygons <- function(x, col) polygon(x[, 1], x[, 2], col = col, density = 10, border = col)
    silent <- mapply(plot.polygons, chull_coords_xy, col_palette)
    legend("bottomlef", legend = LETTERS[1:4], col = unlist(col_palette), bty = "n", pch = 19, lty = 1)

    par(mar = c(4, 4, 1, 3))
    ## Plotting the second coordinates sets
    empty.plot(ylab = "Dimension 3")
    text(-1,1, "2", cex = 2)
    ## Adding the points
    silent <- mapply(plot.points, groups, col_palette, MoreArgs = list(axis = c(1, 3)))
    ## Adding the polygons
    silent <- mapply(plot.polygons, chull_coords_xz, col_palette)
}

plot.figure.dimensionality()

## Figure for the paper
pdf(file = "../Writing/Figures/figure_dimensionality.pdf", width = 5, height = 8.33333)
plot.figure.dimensionality()
dev.off()
```

## Metrics

```{r, fig.width = 10, fig.height = 10}
set.seed(42)

## Scale the results to be between -1 and 1
scale.matrix <- function(matrix) {
    ## Get the range for scaling
    range <- apply(matrix, 2, range)

    ## Shift to positive
    shift_matrix <- matrix
    shift_matrix[,1] <- matrix[,1] + abs(range[1,1])
    shift_matrix[,2] <- matrix[,2] + abs(range[1,2])

    ## Scale the matrix (to + 2)
    shift_matrix <- apply(shift_matrix, 2, function(x) x/(max(x)/2))

    ## Reshift the matrix
    return(shift_matrix - 1)
}

## Plotting the two matrices
plot.both.matrices <- function(matrix1, matrix2, col = c("blue", "orange"), ...) {
    plot(matrix1, pch = 21, bg = col[1], ..., xlab = "", ylab = "")
    points(matrix2, pch = 21, bg = col[2])
}

## Two similar correlation matrices
cor_mat_positive <- matrix(c(1, 0.8, 0.8, 1), 2, 2)
cor_mat_negative <- matrix(c(1, -0.8, -0.8, 1), 2, 2)

## Correlation differences
correlated_positive <- scale.matrix(space.maker(100, 2, rnorm, cor.matrix = cor_mat_positive))
correlated_negative <- scale.matrix(space.maker(100, 2, rnorm, cor.matrix = cor_mat_negative))

## Doughnut differences
outer <- space.maker(100, 2, random.circle, arguments = list(list(runif, inner = 0.5)))
inner <- space.maker(100, 2, random.circle, arguments = list(list(runif, outer = 0.5)))

## Density differences
base_matrix <- space.maker(100, 2, runif, arguments = list(list(max = 1, min = -1)))
modif_matrix <- apply(base_matrix, 2, function(x) x+rnorm(length(x), sd = 0.035))

## Position differences
upper_right <- space.maker(100, 2, rnorm, arguments = list(list(mean = 0.5, sd = 0.2)))
lower_left <- space.maker(100, 2, rnorm, arguments = list(list(mean = -0.5, sd = 0.2)))

plot.figure.metrics <- function(){
    par(mfrow = c(2,2), mar = c(2.5,2.5,2.5,2.5))
    plot.both.matrices(correlated_positive, correlated_negative, main = "Correlation")
    plot.both.matrices(outer, inner, main = "Filling")
    plot.both.matrices(base_matrix, modif_matrix, main = "Density")
    plot.both.matrices(upper_right, lower_left, xlim = c(-1, 1), ylim = c(-1, 1), main = "Position")
}

plot.figure.metrics()

## Figure for the paper
pdf(file = "../Writing/Figures/figure_metrics.pdf", width = 10, height = 10)
plot.figure.metrics()
dev.off()

```

## Figure methods

```{r}
set.seed(42)
## Creating a 5 trait space
iris_rand <- cbind(iris[seq(from = 1, to = 150, by = 5), 1:4], rnorm(30))

## Plotting the 5 traits
pdf(file = "../Writing/Figures/figure_methods_biplot.pdf", width = 5, height = 5)
plot(iris_rand, pch = 19, xaxt = "n", yaxt = "n")
dev.off()

## Making a distance matrix
iris_dist <- dist(iris_rand)

## Making it 2D
pdf(file = "../Writing/Figures/figure_methods_2dPCO.pdf", width = 5, height = 5)
plot(cmdscale(iris_dist, k = 2), pch = 19, xaxt = "n", yaxt = "n", ylab = "", xlab = "")
dev.off()

## Making a PCA
pca <- princomp(iris_rand)$scores
pdf(file = "../Writing/Figures/figure_methods_PCA1.pdf", width = 5, height = 5)
plot(pca[,1:2], pch = 19, xaxt = "n", yaxt = "n", ylab = "", xlab = "")
dev.off()
pdf(file = "../Writing/Figures/figure_methods_PCA2.pdf", width = 5, height = 5)
plot(pca[,3:4], pch = 19, xaxt = "n", yaxt = "n", ylab = "", xlab = "")
dev.off()
pdf(file = "../Writing/Figures/figure_methods_PCA3.pdf", width = 5, height = 5)
plot(pca[,c(5,1)], pch = 19, xaxt = "n", yaxt = "n", ylab = "", xlab = "")
dev.off()
```

## Iris example

```{r}
## Preparing the dataset
colnames(iris)[1:4] <- gsub("\\.", "\n", colnames(iris)[1:4])
rownames(iris) <- c(paste0("setosa", which(iris[,5] == "setosa")),
                    paste0("versicolor", which(iris[,5] == "versicolor")-50),
                    paste0("virginica", which(iris[,5] == "virginica")-100))

## Morphospaces:
## No transformations
write.csv(head(iris[, -5]), file = "../Writing/Figures/trait.space.no_trans.csv")

## Distance matrix
iris_dist <- dist(iris[, -5])
write.csv(round(head(as.matrix(iris_dist)[1:5, 1:5]), digits = 2), file = "../Writing/Figures/trait.space.distance.csv")

## Ordinations
iris_pca <- princomp(iris[, -5])$scores
write.csv(round(head(iris_pca), digits = 2), file = "../Writing/Figures/trait.space.pca.csv")
iris_pco <-cmdscale(iris_dist, k = 2)

## Visualisations
## biplots
col_var <- c("orange", "blue", "green")[as.factor(iris[,5])]
pdf(file = "../Writing/Figures/figure_methods_biplots.pdf", width = 5, height = 5)
plot(iris[, -5],
    pch = 19,
    col = col_var)
dev.off()

## PCO
vars <- round(apply(iris_pco, 2, var)/sum(apply(iris_pco, 2, var))*100, 2)
plot(iris_pco,
    pch = 19,
    col = col_var,
    xlab = paste0("PCO 1 (", vars[1], "%)"),
    ylab = paste0("PCO 21 (", vars[2], "%)"))

## PCA
pdf(file = "../Writing/Figures/figure_methods_PCA.pdf", width = 5, height = 5)
vars <- round(apply(iris_pca, 2, var)/sum(apply(iris_pca, 2, var))*100, 2) 
ranges <- apply(iris_pca, 2, range)

par(mfrow = c(4,4), mar = c(1,1,1,1), oma = c(2,2,2,2))
## First row
# 1
plot(NULL, xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5), xaxt = "n", yaxt = "n", xlab = "", ylab = "")
text(0,0, paste0("PCA 1\n(", vars[1], "%)"), cex = 1)
# 2
plot(iris_pca[, c(1,2)], xaxt = "n", yaxt = "n", ylab = "", xlab = "",
    pch = 19,
    col = col_var)
axis(3)
# 3
plot(iris_pca[, c(1,3)], xaxt = "n", yaxt = "n", ylab = "", xlab = "",
    pch = 19,
    col = col_var)
# 4
plot(iris_pca[, c(1,4)], xaxt = "n", yaxt = "n", ylab = "", xlab = "",
    pch = 19,
    col = col_var)
axis(3)
axis(4)

## Second row
# 1
plot(iris_pca[, c(2,1)], xaxt = "n", yaxt = "n", ylab = "", xlab = "",
    pch = 19,
    col = col_var)
axis(2)
# 2
plot(NULL, xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5), xaxt = "n", yaxt = "n", xlab = "", ylab = "")
text(0,0, paste0("PCA 2\n(", vars[2], "%)"), cex = 1)
# 3
plot(iris_pca[, c(2,3)], xaxt = "n", yaxt = "n", ylab = "", xlab = "",
    pch = 19,
    col = col_var)
# 4
plot(iris_pca[, c(2,4)], xaxt = "n", yaxt = "n", ylab = "", xlab = "",
    pch = 19,
    col = col_var)
axis(4)

## Third row
# 1
plot(iris_pca[, c(3,1)], xaxt = "n", yaxt = "n", ylab = "", xlab = "",
    pch = 19,
    col = col_var)
axis(2)
# 3
plot(iris_pca[, c(3,2)], xaxt = "n", yaxt = "n", ylab = "", xlab = "",
    pch = 19,
    col = col_var)
# 3
plot(NULL, xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5), xaxt = "n", yaxt = "n", xlab = "", ylab = "")
text(0,0, paste0("PCA 3\n(", vars[3], "%)"), cex = 1)
# 4
plot(iris_pca[, c(3,4)], xaxt = "n", yaxt = "n", ylab = "", xlab = "",
    pch = 19,
    col = col_var)
axis(4)

## Fourth row
# 1
plot(iris_pca[, c(4,1)], xaxt = "n", yaxt = "n", ylab = "", xlab = "",
    pch = 19,
    col = col_var)
axis(1)
axis(2)
# 3
plot(iris_pca[, c(4,2)], xaxt = "n", yaxt = "n", ylab = "", xlab = "",
    pch = 19,
    col = col_var)
# 3
plot(iris_pca[, c(4,3)], xaxt = "n", yaxt = "n", ylab = "", xlab = "",
    pch = 19,
    col = col_var)
axis(1)
# 4
plot(NULL, xlim = c(-0.5, 0.5), ylim = c(-0.5, 0.5), xaxt = "n", yaxt = "n", xlab = "", ylab = "")
text(0,0, paste0("PCA 4\n(", vars[4], "%)"), cex = 1)
dev.off()