### 2020-03-13 Writing/CoverLetter.tex
cover letter  INIT

\documentclass[11pt]{letter}
\usepackage[a4paper,left=2.5cm, right=2.5cm, top=1cm, bottom=1cm]{geometry}
\usepackage{hyperref}
\usepackage[osf]{mathpazo}
\signature{Thomas Guillerme, on behalf of my co-authors.}
\address{The University of Queensland \\School of Biological Sciences \\St Lucia QLD 4067, Australia \\guillert@tcd.ie}
\longindentation=0pt
\begin{document}

\begin{letter}{}
\opening{Dear Editors,}

Royal Society International Scientific Seminar: Reconciling disparate perspectives on the evolution of disparity on the 9th and 10th of January 2018 in Chicheley Hall.

Along with all 19 the meeting participants and 3 additional researchers, we wrote a review entitled: ``Disparities in the analysis of morphological disparity''.

This paper covers:
- The origin of disparity in litterature
- The types of analysis
- The types of data
- The types of statistical analysis (including phylogeny)
- The future of disparity methods

This paper especially aims to provide a common framework to disparity analysis that can be used as a base for macroevolution researchers to make choices in their analysis and go further.

It also aims to foster interdisciplinarity and disparity beyond morphology (e.g. in ecology or isotopes analysis).

Overall, this broad review proposes the state-of-the-art in disparity analyses and where the field has been going since its advent in the 1980s.

We look forward to hearing from you soon,

\closing{Yours sincerely,}

\end{letter}
\end{document}

%Editors:
%Sam Price, Daniele Silvestro

%Reviewers
%Graeme Lloyd, Dave Bapst, Emma Sherratt